Properties of Winning Iterated Prisoner's Dilemma Strategies.

### 2020-04-14 paper/main.tex
reads over evaluation of performance
Fixes correlation table

The memory usage of strategies is the number of
rounds of play used by the strategy divided by the number of turns in each match. 
For example, Winner12 uses the previous two rounds of play, and if participating in a
participated in a match with 100 turns, the memory usage would be 2/100.
For strategies with an infinite memory size, for example Evolved
FSM 16 Noise 05, is equal to 1.
Note that for tournaments with a probabilistic
ending the number of turns was not collected, so the memory usage feature is not
used for probabilistic ending tournaments.

@@

Moreover, the manner in which a strategy achieves a given cooperation rate
relative to the tournament population average is important. Playing a strategy
that randomly cooperates with $C_{\text{mean}}$ is unlikely to be as effective as it will
randomly have cooperations matched with defections, and it may be punished for
defections against cooperations by retailiating opponents. In contrast, TFT naturally
achieves a cooperation rate near $C_{\text{mean}}$ by virtue of copying its opponent's
last move while also minimizing instances where it is exploited by an opponent
(cooperating while the opponent defects), at least in non-noisy tournaments.
\footnote{This also explains why Tit For N Tats does not fare well for $N > 1$ -- it fails
to achieve the proper cooperation ratio by tolerating too many defections.}
TFT forces the opponent to pay back the (C, D) round with a (D, C) round before
returning to mutual cooperation. This explains why TFT performed well
in Axelrod's original tournaments as most strategies submitted to those tournaments
were typically cooperative and relatively few strategies played with an easily exploitable
pattern. TFT does not appear in the top ranks in tournaments with more aggressive
strategies because it is too nice. Strategies like Grudger will always
defect after a fixed number of opponent defections, which allows them to
effectively exploit strategies like Alternator or stochastic strategies that
have a non-zero chance of cooperating after mutual defection, which TFT will
not do. Moreover in a noisy environment these strategies will naturally tend
toward always defecting, leading them to exploit strategies like Cooperator. In
such a scenario the noisy environment effectively voids Axelrod's rule to be nice,
allowing strategies to attempt exploitation, whereas in a noise-free environment,
attempting exploitation can be risky if strategies that exhibit a Grudger-like behavior
are present, reducing the overall value in attempting to exploit strategies like Cooperator.

Similarly, these results suggest an explanation regarding the intuitively unexpected
effectiveness of memory one strategies historically. Given that among the important
features associated with success are the relative cooperation rate to the population
average and the four memory-one probabilities of cooperating conditional on the
previous round of play, these features can be optimized by a memory one
strategy such as TFT. Usage of more history becomes valuable when there are
exploitable opponent patterns. This is indicated by the importance of SSE as a feature,
showing that the first-approximation provided by a memory one strategy is no longer
sufficient.

@@

The distributions of SSE and \(CC\) to \(D\) rate for the winners of standard
tournaments are also given in Figure~\ref{fig:discussion_standard}. The SSE
distributions for the winners indicate that the strategy behaved in a ZD way in
several tournaments, however, not constantly. The winners participated in
matches where they did not try to extortionate their opponents. Furthermore, the
\(CC\) to \(D\) distribution indicates that if a strategy were to defect against
the winners they would reciprocate on average with a probability of 0.5.

>>>>

The memory usage of strategies is the number of
rounds of play used by the strategy divided by the number of turns in each match.
For example, Winner12 uses the previous two rounds of play, and if participating
in a match with 100 turns its memory usage would be 2/100.
For strategies with an infinite memory size, for example Evolved
FSM 16 Noise 05, memory usage is equal to 1.
Note that for tournaments with a probabilistic
ending the number of turns was not collected, so the memory usage feature is not
used for probabilistic ending tournaments.

@@

Moreover, the manner in which a strategy achieves a given cooperation rate
relative to the tournament population average is important. Playing a strategy
that randomly cooperates with $C_{\text{mean}}$ is unlikely to be as effective as it will
randomly have cooperations matched with defections, and it may be punished for
defections against cooperations by retailiating opponents. In contrast, TFT naturally
achieves a cooperation rate near $C_{\text{mean}}$ by virtue of copying its opponent's
last move while also minimizing instances where it is exploited by an opponent
(cooperating while the opponent defects), at least in non-noisy tournaments.
\footnote{This also explains why Tit For N Tats does not fare well for $N > 1$ -- it fails
to achieve the proper cooperation ratio by tolerating too many defections.}
TFT forces the opponent to pay back the \((C, D)\) round with a \((D, C)\) round before
returning to mutual cooperation. This explains why TFT performed well
in Axelrod's original tournaments as most strategies submitted to those tournaments
were typically cooperative and relatively few strategies played with an easily exploitable
pattern. TFT does not appear in the top ranks in tournaments with more aggressive
strategies because it is too nice. Strategies like Grudger will always
defect after a fixed number of opponent defections, which allows them to
effectively exploit strategies like Alternator or stochastic strategies that
have a non-zero chance of cooperating after mutual defection, which TFT will
not do. Moreover in a noisy environment these strategies will naturally tend
toward always defecting, leading them to exploit strategies like Cooperator. In
such a scenario the noisy environment effectively voids R. Axelrod's rule to be nice,
allowing strategies to attempt exploitation, whereas in a noise-free environment,
attempting exploitation can be risky if strategies that exhibit a Grudger-like behavior
are present, reducing the overall value in attempting to exploit strategies like Cooperator.

Similarly, these results suggest an explanation regarding the intuitively unexpected
effectiveness of memory-one strategies historically. Given that among the important
features associated with success are the relative cooperation rate to the population
average and the four memory-one probabilities of cooperating conditional on the
previous round of play, these features can be optimized by a memory-one
strategy such as TFT. Usage of more history becomes valuable when there are
exploitable opponent patterns. This is indicated by the importance of SSE as a feature,
showing that the first-approximation provided by a memory-one strategy is no longer
sufficient.

@@

The distributions of SSE and \(CD\) to \(D\) rate for the winners of standard
tournaments are also given in Figure~\ref{fig:discussion_standard}. The SSE
distributions for the winners indicate that the strategy behaved in a ZD way in
several tournaments, however, not constantly. The winners participated in
matches where they did not try to extortionate their opponents. Furthermore, the
\(CD\) to \(D\) distribution indicates that if a strategy were to defect against
the winners they would reciprocate on average with a probability of 0.5.
